Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-04-07 docs/ACM/paper.tex
Adding related work session + new references.

>>>>

Data-flow programming was originally proposed as a candidate to enable
task-based parallelism \cite{agrawal:ipdps:2010, chan:spaa:2007,
gupta:micro:2011}.  Agrawal et al. \cite{agrawal:ipdps:2010}  extended
data-flow programming  with task input/otuput  specifications in  Cilk++.
Vandierendonck et al.  \cite{vandierendonck:hotpar:2011} also extended Cilk++
by adding dependency clauses that facilitated the design of complex
parallelization patterns. Both works presented a unified scheduler based on
fork-join parallelism \cite{vandierendonck:pact:2011} which enabled the
execution of task-based dataflow applications. Other approaches have also used
data-flow task graphs to exploit parallelism, like Data-Driven Tasks
\cite{tasirlar:icpp:2011}, where the programmer can use put/await clauses to
determine the task arguments before execution.  These works focused on
data-flow programming  models and unfortunately  have not been adopted by most
programmers. Function-based task parallelism was proposed by Gupta et al
\cite{gupta:micro:2011} with the goal of using function arguments  as a way to
specify task  dependencies. Although their work seemed sound, programming using
such approach is a cumbersome task, when compared to modern directive-based
parallelization.

Many systems have been developed aiming at (semi) automatically extracting task
parallelism from sequential programs. Examples of such systems are OSCAR
(Optimally SCheduled Advanced multiprocessoR), Multigrain Parallelizing
Compiler \cite{ishizaka:journal:2000, kasahara:iwlcpc:2000},  MAPS (MPSoC
Application Programming  Studio) \cite{castrillon:tii:2013, ceng:dac:2008}, and
DiscoPoP (Discovery of Potential Parallelism) \cite{discopop, li:jss:2016}.
Although most of these systems have succeeded in achieving their goals, overall
they have not managed to engage most programmers in using them. Besides
parallelization systems, tools like Paraver \cite{extrae, paraver}, Aftermath
\cite{drebes:hipeac:2014}, DAGvis \cite{huynh:wvpa:2015}, and TEMANEJO
\cite{brinkmann:parco:2011, brinkmann:journal:2013, temanejo} were  designed to
enable performance analysis, debugging, and vizualization of task programming
executions.

More recently, directive based sequential code annotation  standards have
provided programmers with a very clean and easy to use parallel programming
model. Task based extensions to these standards have resulted in OpenMP 4.X,
StarSs \cite{bellens:sc:2006, perez:iccc:2008, planas:hpca:2009,
tejedor:hpdc:2011} and OmpSs \cite{bueno:icpp:2011, duran:ppl:2011}  which have
been extensively used to task-parallelize a large number of applications. Such
programming models have the support of tools that help programmers find the
best annotations for their code. For example,  Tareador
\cite{ayguade:wcae:2015} enables a programmer, by means of a simple API and a
graphical representation,  to iteratively annotate sequential code, thus
allowing the identification of potential task parallelization opportunities.
Contrary to Tareador,  this paper proposes an approach that enables the
automatic insertion OpenMP task annotations to relevant fragments of a
sequential program.
